### 2020-02-01 tables/datastats.tex

% \begin{table}[!t]
    % \centering
    % % \small{
    % \caption{Data statistics. The abbreviations for subgroups used throughout report.}
    \begin{tabular}{p{2.25cm} m{1cm}>{\centering} p{1.2cm}  m{1cm}>{\centering}} % Columns will widen to hold the content in them by default, to specify a precise width, use p{width}, e.g. p{5cm}
        % \toprule 
        % & \multicolumn{5}{c}{\textbf{Subgroups}} \\
        % \cmidrule(l){2-6} % Horizontal line spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
       Subgroup &\# Faces &\# Subjects &\# Pairs\\ % Column names row
        \midrule
        \gls{af}&1,997 & 100   &52,980 \\ % row 1
        \gls{am}&2,002 & 100   &50,470  \\ % row 2
        \gls{bf}& 1,763& 100   &56,190  \\ % row 3
        \gls{bm}& 1,803&   100  &55,805 \\ % row 4
        \gls{if}& 1,902 &  100  &38,265 \\ % row 5
        \gls{im}& 1,890&  100   &41,560 \\ % row 6
        \gls{wf}&1,869 &  100   &59,705 \\ % row 7
        \gls{wm}&1,969 &  100   &58,715 \\ % row 819150 pairs for asian_females

        \midrule
        % \midrule
        Total & 15,195 &  799    & 413,690 \\ % Summary/total row
        % \bottomrule
    \end{tabular}
    % }
    % \label{tbl:datastats} 
% \end{table}