### 2020-02-01 tables/gender_splits.tex

\begin{tabular}{l c c c c} % Columns will widen to hold the content in them by default, to specify a precise width, use p{width}, e.g. p{5cm}
    \toprule 
    & \multicolumn{2}{c}{\textbf{Females}} & \multicolumn{2}{c}{\textbf{Males}}        \\
    \cmidrule(l){2-3} \cmidrule(l){4-5} % Horizontal line spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
     & A & B & A & B\\ % Column names row
    \midrule
    No. Positive & 28,354 & 28,354 & 29,479 & 29,479\\ % row 1
    No. Negative &39,970 & 129,389 & 53,535& 170,047 \\ % row 2
    Total & 68,324 & 157,743 &  83,014&199,526 \\ % row 3
    \midrule
    Threshold & 0.191 & 0.317 & 0.187 & 0.300 \\ % row 4
    Top Acc. & 0.9175 & 0.9287 & 0.9180 &0.9369 \\ % row 5
    % \bottomrule
\end{tabular}