### 2020-02-01 tmp.tex

\begin{tabular}{lllr}\\\toprule\\ &   &    &     p1 \\ fold & label & type &        \\\midrule\\ 1 & 0 & bb &  34729 \\ &   & fd &  16605 \\ &   & fs &  27597 \\ &   & gfgd &    604 \\ &   & gfgs &    736 \\ &   & gmgd &    621 \\ &   & gmgs &    726 \\ &   & md &  13372 \\ &   & ms &  14996 \\ &   & sibs &  13344 \\ &   & ss &  34728 \\ & 1 & bb &  29275 \\ &   & fd &  13596 \\ &   & fs &  28108 \\ &   & gfgd &    916 \\ &   & gfgs &    854 \\ &   & gmgd &    711 \\ &   & gmgs &    565 \\ &   & md &  12924 \\ &   & ms &  32597 \\ &   & sibs &   8765 \\ &   & ss &  29275 \n2 & 0 & bb &  18880 \\ &   & fd &   7508 \\ &   & fs &   8461 \\ &   & gfgd &    401 \\ &   & gfgs &    403 \\ &   & gmgd &    829 \\ &   & gmgs &    848 \\ &   & md &   6915 \\ &   & ms &  10638 \\ &   & sibs &   6681 \\ &   & ss &  18882 \\ & 1 & bb &  28476 \\ &   & fd &   7888 \\ &   & fs &  12338 \\ &   & gfgd &   1181 \\ &   & gfgs &    309 \\ &   & gmgd &    960 \\ &   & gmgs &    407 \\ &   & md &   8474 \\ &   & ms &   9022 \\ &   & sibs &  10342 \\ &   & ss &  28476 \\3 & 0 & bb &  23981 \\ &   & fd &   9744 \\ &   & fs &  12237 \\ &   & gfgd &   1456 \\ &   & gfgs &   1014 \\ &   & gmgd &    391 \\ &   & gmgs &    386 \\ &   & md &   9314 \\ &   & ms &  15235 \\ &   & sibs &   7143 \\ &   & ss &  23997 \\ & 1 & bb &  19363 \\ &   & fd &   9369 \\ &   & fs &  11039 \\ &   & gfgd &    829 \\ &   & gfgs &    772 \\ &   & gmgd &    583 \\ &   & gmgs &    609 \\ &   & md &   8850 \\ &   & ms &   8937 \\ &   & sibs &   5264 \\ &   & ss &  19363 \\ 4 & 0 & bb &  14016 \\ &   & fd &   7262 \\ &   & fs &   9371 \\ &   & gfgd &    913 \\ &   & gfgs &    766 \\ &   & gmgd &    948 \\ &   & gmgs &    593 \\ &   & md &   8089 \\ &   & ms &  11026 \\ &   & sibs &   5149 \\ &   & ss &  14013 \\ & 1 & bb &  15838 \\ &   & fd &   7601 \\ &   & fs &   7202 \\ &   & gfgd &    337 \\ &   & gfgs &    662 \\ &   & gmgd &    269 \\ &   & gmgs &    552 \\ &   & md &   8367 \\ &   & ms &   6565 \\ &   & sibs &   7182 \\ &   & ss &  15838 \\ 5 & 0 & bb &  18671 \\ &   & fd &   8733 \\ &   & fs &  11421 \\ &   & gfgd &    690 \\ &   & gfgs &    431 \\ &   & gmgd &    664 \\ &   & gmgs &    381 \\ &   & md &   8970 \\ &   & ms &  12984 \\ &   & sibs &   7477 \\ &   & ss &  18661 \\ & 1 & bb &  17310 \\ &   & fd &  11385 \\ &   & fs &  10393 \\ &   & gfgd &    803 \\ &   & gfgs &    753 \\ &   & gmgd &    930 \\ &   & gmgs &    803 \\ &   & md &   8058 \\ &   & ms &   7768 \\ &   & sibs &   8238 \\ &   & ss &  17310 \\\bottomrule
\end{tabular}